%%%%
% Consiglio la visione dei seguenti tutorial:
% - https://www.youtube.com/watch?v=ihxSUsJB_14
% - https://www.youtube.com/watch?v=XTFWaV55uDo
%%%%
\documentclass[12pt,a4paper,openright,twoside]{book} % TODO: togliere draft
\usepackage[utf8]{inputenc}

%\newcommand{\thesislang}{italian} % Italian thesis
\newcommand{\thesislang}{english} % English thesis
\usepackage{thesis-style}
% version
\newcommand{\versionmajor}{0}
\newcommand{\versionminor}{1}
\newcommand{\versionpatch}{2}
\newcommand{\version}{\versionmajor.\versionminor.\versionpatch}
\typeout{Document version: \version}

\begin{document}

\frontmatter

\input{front.tex}

\begin{abstract}
	The emergence of cyber-physical systems has brought about a significant increase in complexity and heterogeneity in the infrastructure on which
	these systems are deployed. One particular example of this complexity is the interplay between cloud, fog, and edge computing. However, the
	complexity of these systems can pose challenges when it comes to implementing self-organizing mechanisms, which are often designed to work on
	flat networks. Therefore, it is essential to separate the application logic from the specific deployment aspects to promote reusability and
	flexibility in infrastructure exploitation.

	To address this issue, a novel approach called "pulverization" has been proposed. This approach involves breaking down the system into smaller
	computational units, which can then be deployed on the available infrastructure. In this thesis, the design and implementation of a portable
	framework that enables the ``pulverization'' of cyber-physical systems are presented.
	The main objective of the framework is to pave the way for the deployment of cyber-physical systems in the edge-cloud continuum
	by reducing the complexity of the infrastructure and exploit opportunistically the heterogeneous resources available on it. Different scenarios
	are presented to highlight the effectiveness of the framework in different heterogeneous infrastructures and devices. Current limitations and
	future work are examined to identify improvement areas for the framework.
\end{abstract}

\begin{dedication} % this is optional
	Alla mia famiglia.
\end{dedication}

\begin{acknowledgements} % this is optional
	First of all, I would like to express my gratitude towards Professors Viroli and Pianini for their invaluable assistance and insightful guidance,
	which was essential for the realization of this thesis work.
	%
	I thank my family infinitely for their continuous support and motivation during these years of study and for never letting me lack anything.
	%
	A special thanks go to Giulia who has always been there in these two years supporting and putting up with me in both the most difficult and the
	most beautiful moments of this path.
	%
	I would also like to extend my thanks to the members of the Atedeg group with whom I have shared countless unforgettable moments and experiences.
	Additionally, I am grateful to all my university colleagues for the enjoyable times we have spent together.
	%
	Lastly, but certainly not least, I am grateful to all my friends who have made these two important years so joyful and carefree. To everyone who
	believed in me and supported me during this time, I express my heartfelt thanks.
\end{acknowledgements}

%----------------------------------------------------------------------------------------
\tableofcontents
\listoffigures     % (optional) comment if empty
\lstlistoflistings % (optional) comment if empty
%----------------------------------------------------------------------------------------

\mainmatter

% Introduction --------------------------------------------------------------------------
\input{chapters/1-introduction.tex}
%----------------------------------------------------------------------------------------

% State of the art ----------------------------------------------------------------------
\input{chapters/2-state-of-art.tex}
%----------------------------------------------------------------------------------------

% Requirements --------------------------------------------------------------------------
\input{chapters/3-requirements.tex}
%----------------------------------------------------------------------------------------

% Design --------------------------------------------------------------------------------
\input{chapters/4-design.tex}
%----------------------------------------------------------------------------------------

% Implementations -----------------------------------------------------------------------
\input{chapters/5-implementation.tex}
%----------------------------------------------------------------------------------------

% Validations ---------------------------------------------------------------------------
\input{chapters/6-validation.tex}
%----------------------------------------------------------------------------------------

% Conclusions ---------------------------------------------------------------------------
\input{chapters/7-conclusions.tex}
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
% BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

%\nocite{*} % uncomment this to show all the references in the .bib file
% \bibliographystyle{plain}
% \bibliography{bibliography}
\printbibliography

\end{document}
